\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-17 paper/paper 2/dc-2015.tex
introduction

\maketitle
\vspace{-3em}
\section*{Abstract}
In the context of the DCMI RDF Application Profile task group and the W3C Data Shapes Working Group solutions for the proper formulation of constraints and validation of RDF data on these constraints are developed. Several approaches and constraint languages exist but there is no clear favorite and none of the languages is able to meet all requirements raised by data practitioners.
To support the work, a comprehensive, community-driven database has been created where case studies, use cases, requirements and solutions are collected. Based on this database,
we published by today 81 types of constraints that are required by various stakeholders for data applications. We generally use this collection of constraint types to gain a better understanding of the expressiveness of existing solutions and gaps that still need to be filled. 
Regarding the implementation of constraint languages, we already proposed to use high-level languages to describe the constraints, but map them to SPARQL queries in order to execute the actual validation; we demonstrated this approach for \textcolor{red}{the Web Ontology Language in its current version 2 and }Description Set Profiles.
In this paper, we generalize from the experience of implementing \textcolor{red}{OWL 2 and} DSP by introducing an abstraction layer that is able to describe \textcolor{red}{constraints of} any constraint type in a way that mappings from high-level constraint languages to this intermediate representation can be created more or less straight-forwardly. 
We demonstrate that using another layer on top of SPARQL helps to implement validation consistently accross constraint languages, simplifies the actual implementation of new languages, and supports the transformation of semantically equivalent constraints across constraint languages.

@@

Various constraint languages exist or are being developed that support more or less of these constraint types. For our work, we focus on the following four as the ones that are most popular among data practitioners, often mentioned on mailing lists and/or being candidates or prototypes for the upcoming W3C recommendation:
\emph{Description Set Profiles (DSP)}\footnote{\url{ http://dublincore.org/documents/2008/03/31/dc-dsp/}}, \emph{Resource Shapes (ReSh)}\footnote{\url{http://www.w3.org/Submission/2014/SUBM-shapes-20140211/}}, \emph{Shape Expressions (ShEx)}\footnote{\url{http://www.w3.org/Submission/2014/SUBM-shex-primer-20140602/}}, and the \emph{Web Ontology Language} \emph{(OWL)}\footnote{\url{http://www.w3.org/TR/owl2-syntax/}}. Despite the fact that OWL is arguably not a constraint language, it is widely used in practice as such under the closed-world and unique name assumptions.

@@

The creation of such mappings is in many cases not straight-forward and requires profound knowledge of SPARQL, as the following example demonstrates. 
The constraint type \emph{minimum qualified cardinality restrictions} which corresponds to the requirement \emph{R-75}\footnote{Requirements are identified in the database by an R and a number, additionally an alphanumeric identifier is provided, in this case R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES. Online at: \url{http://lelystad.informatik.uni-mannheim.de/rdf-validation/?q=node/82}} can be used to formulate the constraint that publications must have at least one author which must be a person. This constraint can be expressed as follows using different constraint languages:

@@

\textit{DSP:} [ dsp:resourceClass Publication ; dsp:statementTemplate [ 
          dsp:minOccur 1 ; dsp:maxOccur "infinity" ; 
          dsp:property author ; 
          dsp:nonLiteralConstraint [ dsp:valueClass Person ] ] ] .

@@

@@

\textit{DSP:} [ dsp:resourceClass Publication ; dsp:statementTemplate [ 
          dsp:minOccur 1 ; dsp:maxOccur "infinity" ; 
          dsp:property author ; 
          dsp:nonLiteralConstraint [ dsp:valueClass Person ] ] ] .
\end{ex}

@@

At the core, we use a very simple conceptual model for constraints, using a small vocabulary called \emph{RDF Constraints Vocabulary (RDF-CV)}\footnote{Online available at: \url{https://github.com/boschthomas/RDF-Constraints-Vocabulary}}. The conceptual model is shown in Figure \ref{fig:RDF-CV-conceptual-model}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.60\textwidth]{images/RDF-CV.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Conceptual Model}
	\label{fig:RDF-CV-conceptual-model}
\end{figure}

@@

Finally, the \emph{constraining value} contains a literal value to be checked against; for instance in the case of the 'REGEX' element, it contains the regular expression to be evaluated.

@@

The mapping simply constructs this representation out of the OWL2 representation using a SPARQL CONSTRUCT query:

@@

As stated in Section~\ref{sec:motivation}, we see a huge potential in the possibility to transform semantically equivalent constraints from one high-level constraint language to another via the RDF-CV representation, to avoid that every possible combination of constraint languages has to be mapped separately. The following SPIN inference rule exemplifies this approach and provides a mapping from RDF-CV back to an OWL2 constraint:

@@

\tb{ToDo Thomas: OWL 2}
\emph{Stardog Integrity Constraint Validation (ICV)} and the \emph{Pellet Integrity Constraint Validator (ICV)} use OWL 2 constructs to formulate constraints. The Pellet ICV\footnote{\url{http://clarkparsia.com/pellet/icv}} is a proof-of-concept extension for the OWL 2 DL reasoner \emph{Pellet} \cite{sirin2007pellet}. Stardog ICV\footnote{\url{http://docs.stardog.com/#_validating_constraints}} validates RDF data stored in a Stardog database according to constraints which may be written in SPARQL, OWL 2, or SWRL \cite{Horrocks04}.

>>>>

\maketitle
\vspace{-3em}
\section*{Abstract}
In the context of the DCMI RDF Application Profile task group and the W3C Data Shapes Working Group solutions for the proper formulation of constraints and validation of RDF data on these constraints are developed. Several approaches and constraint languages exist but there is no clear favorite and none of the languages is able to meet all requirements raised by data practitioners.
To support the work, a comprehensive, community-driven database has been created where case studies, use cases, requirements and solutions are collected. Based on this database,
we published by today 81 types of constraints that are required by various stakeholders for data applications. We generally use this collection of constraint types to gain a better understanding of the expressiveness of existing solutions and gaps that still need to be filled. 
Regarding the implementation of constraint languages, we already proposed to use high-level languages to describe the constraints, but map them to SPARQL queries in order to execute the actual validation; we demonstrated this approach for the Web Ontology Language in its current version 2 and Description Set Profiles.
In this paper, we generalize from the experience of implementing OWL 2 and DSP by introducing an abstraction layer that is able to describe constraints of any constraint type in a way that mappings from high-level constraint languages to this intermediate representation can be created more or less straight-forwardly. 
We demonstrate that using another layer on top of SPARQL helps to implement validation consistently accross constraint languages, simplifies the actual implementation of new languages, and supports the transformation of semantically equivalent constraints across constraint languages.

@@

Various constraint languages exist or are being developed that support more or less of these constraint types. For our work, we focus on the following four as the ones that are most popular among data practitioners, often mentioned on mailing lists and/or being candidates or prototypes for the upcoming W3C recommendation:
\emph{Description Set Profiles (DSP)}\footnote{\url{http://dublincore.org/documents/2008/03/31/dc-dsp/}}, \emph{Resource Shapes (ReSh)}\footnote{\url{http://www.w3.org/Submission/2014/SUBM-shapes-20140211/}}, \emph{Shape Expressions (ShEx)}\footnote{\url{http://www.w3.org/Submission/2014/SUBM-shex-primer-20140602/}}, and the \emph{Web Ontology Language} \emph{(OWL)}\footnote{\url{http://www.w3.org/TR/owl2-syntax/}}. Despite the fact that OWL is arguably not a constraint language, it is widely used in practice as such under the closed-world and unique name assumptions.

@@

The constraint type \emph{minimum qualified cardinality restrictions} which corresponds to the requirement \emph{R-75}\footnote{Requirements are identified in the database by an R and a number, additionally an alphanumeric identifier is provided, in this case R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES. Online at: \url{http://lelystad.informatik.uni-mannheim.de/rdf-validation/?q=node/82}} can be used to formulate the constraint that publications must have at least one author which must be a person. This constraint can be expressed as follows using different constraint languages:

@@

\textit{DSP:} [ dsp:resourceClass Publication ; dsp:statementTemplate [ 
          dsp:minOccur 1 ; 
          dsp:property author ; 
          dsp:nonLiteralConstraint [ dsp:valueClass Person ] ] ] .

@@

The creation of mappings of constraint languages to SPIN is in many cases not straight-forward and requires profound knowledge of SPARQL, as the following example demonstrates. In this example, the validation of the \emph{minimum qualified cardinality restrictions} constraint type is implemented for DSP:

\begin{ex}[commandchars=\\\{\}]
CONSTRUCT \{
    _:constraintViolation 
        a spin:ConstraintViolation ;
        rdfs:label ?violationMessage ;
        spin:violationRoot ?this ;
        spin:violationPath ?property ;
        spin:violationSource ?violationSource . \}
WHERE \{	
    ?this a ?resourceClass .
    ?descriptionTemplate 
        dsp:resourceClass ?resourceClass ;
        dsp:statementTemplate ?statementTemplate .
    ?statementTemplate 
        dsp:minOccur ?minimum ;
        dsp:property ?property ;
        dsp:nonLiteralConstraint ?nonLiteralConstraint .
    ?nonLiteralConstraint dsp:valueClass ?valueClass .	
    BIND ( \textit{qualifiedCardinality}( ?this, ?property, ?valueClass ) AS ?cardinality ) .
    FILTER ( ?cardinality < ?minimum ) . \}
\end{ex}

@@

\textit{DSP:} [ dsp:resourceClass Publication ; dsp:statementTemplate [ 
          dsp:minOccur 1 ; 
          dsp:property author ; 
          dsp:nonLiteralConstraint [ dsp:valueClass Person ] ] ] .
\end{ex}

@@

At the core, we use a very simple conceptual model for constraints, using a small vocabulary called \emph{RDF Constraints Vocabulary (RDF-CV)}\footnote{Online available at: \url{https://github.com/boschthomas/RDF-Constraints-Vocabulary}}. The conceptual model is shown in Figure \ref{fig:RDF-CV-conceptual-model}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CV.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Conceptual Model}
	\label{fig:RDF-CV-conceptual-model}
\end{figure}

@@

Finally, the \emph{constraining value} contains a literal value to be checked against; for instance in the case of the REGEX element, it contains the regular expression to be evaluated.

@@

The mapping simply constructs this representation out of the OWL 2 representation using a SPARQL CONSTRUCT query:

@@

As stated in Section~\ref{sec:motivation}, we see a huge potential in the possibility to transform semantically equivalent constraints from one high-level constraint language to another via the RDF-CV representation, to avoid that every possible combination of constraint languages has to be mapped separately. The following SPIN inference rule exemplifies this approach and provides a mapping from RDF-CV back to an OWL 2 constraint:

@@

\emph{Stardog Integrity Constraint Validation (ICV)} and the \emph{Pellet Integrity Constraint Validator (ICV)} use OWL 2 constructs to formulate constraints. The Pellet ICV\footnote{\url{http://clarkparsia.com/pellet/icv}} is a proof-of-concept extension for the OWL 2 DL reasoner \emph{Pellet} \cite{sirin2007pellet}. Stardog ICV\footnote{\url{http://docs.stardog.com/#_validating_constraints}} validates RDF data stored in a Stardog database according to constraints which may be written in SPARQL, OWL 2, or SWRL \cite{Horrocks04}.
